---
title: "ps_04"
author: "Albert Shin"
date: "2/25/2019"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
# tidyverse for necessary functions like mutate
library(tidyverse)
# package for read_csv
library(readr)
#package for creating a table
library(gt)
# package for creating a visual graph
library(ggplot2)
```

```{r madlibs, echo=FALSE, message=FALSE}
elections = read_csv("ps_4_elections-poll-nc09-3.csv")

#There were X respondents who supported the Democratic candidate.
# assign q1 to the value that we get
q1 <-elections %>% 
  # filter for responses that were "Dem". Used response == "Dem" because we are looking at respondents in particular.
  filter(response == "Dem") %>%
  # count the filtered responses
  count()

#There were X more respondents who favored the Republican candidate than who were Undecided.
# assign the value we get to repfavor, to find the number of people who responded "Rep"
repfavor <- elections %>% 
  # select response (only variable we need)
  select(response) %>% 
  # filter for the responses that responded "Rep"
  filter(response == "Rep") %>% 
  # count the filtered responses
  count()

# assign the value we get to undecided, for the people who responded undecided
undecided <- elections %>% 
  # select response (only variable we need)
  select(response) %>% 
  # filter for the responses that responded "Und"
  filter(response == "Und") %>% 
  # count the filtered responses
  count()

# assign q2 to the difference between repfavor[1] and undecided[1]. We use the bracket 1 ([1]) because we are finding the first observation of repfavor and undecided
q2 <- repfavor[1] - undecided[1]

# There are two gender variables (gender and gender_combined). 
#There are X individuals for whom these variables have different values.
# assign the value we get to q3
q3 <- elections %>% 
  # select for gender and gender_combined, because we are trying to find people who have different values for these
  select(gender, gender_combined) %>% 
  # filter where the value of gender does not equal the value of gender combined
  filter(gender != gender_combined) %>% 
  # count these values
  count()

#There are X respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.
# assign the value we get to q4
q4 <- elections %>% 
  # select race_eth and file_race_black
  select(race_eth, file_race_black) %>% 
  # filter where race_ethnicity equals "White" because the first part of the problem is "X respondents listed as 'white'"
  filter(race_eth == "White") %>% 
  # filter where file_race_black is not listed as white, out of those whose race_eth is "White"
  filter(file_race_black != "White") %>% 
  # count these values
  count()

#The first response of Rep came X minutes (rounded to the nearest minute) before the first response of Dem.
# find the earliest time of Dem 
dem_time <- elections %>% 
  # filters for only dem responses
  filter(response == "Dem") %>% 
  # pulls out a single observation value
  pull(timestamp)

# find the earliest time of Rep 
rep_time <- elections %>% 
  # filters for only rep responses
  filter(response == "Rep") %>% 
  # pulls out a single observation value
  pull(timestamp)

# assign q5 to the difference in timestamps between reptime and demtime
q5 <-round(rep_time[1] - dem_time[1])
```
##Question 1
<!-- answering the madlibs; used dynamic inline coding to produce the answer on the document --> 
There were `r q1` respondents who supported the Democratic candidate.

There were `r q2` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (gender and gender_combined). There are `r q3` individuals for whom these variables have different values.

There are `r q4` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

The first response of Dem came `r q5` minutes (rounded to the nearest minute) before the first response of Rep.

##Question 2
```{r table, echo=FALSE, warning=FALSE}
table <- elections %>%
  # select the variables we need: race/ethnicity, and final weight (for the percentage of total voters)
  select(response, race_eth, final_weight) %>%
  # Take out the entries who refused to say / don't know their ethnicity 
  filter(race_eth != "[DO NOT READ] Don't know/Refused") %>%
  # Use mutate in order to reorder the ethnicities in the correct order
  mutate(race_eth = fct_relevel(race_eth, "White", "Black", "Hispanic", "Asian", "Other")) %>%
  # group by response and then by race_eth to help calculate the individual table values
  group_by(response, race_eth) %>%
  # total equals the sum of the final weights - we add up the final weights because the totals of that are equivalent to the total amount of votes. Published results include weights 
  summarize(total = sum(final_weight)) %>%
  # reformats table. Takes the key(response) and value (total) and spreads them over multiple columns
  spread(key = response, value = total, fill = 0) %>%
  # changes total to sum of Dem responses + Rep responses + Und responses + third party responses
  mutate(total = Dem + Rep + Und + `3`) %>%
  # muates Dem and Rep and Und by dividing it by the total value (which is the sum of dem and rep and und and 3, so the total number of responses) 
  mutate(Dem = Dem / total) %>%
  mutate(Rep = Rep / total) %>%
  mutate(Und = Und / total) %>%
  # selects everything except for total and `3`, because we do not need those values 
  select(-total, -`3`)

gt(table) %>%
  tab_header(
    title = "Polling Results by Race in North Carolina's 9th Congressional District") %>% 
  # Labels the columns (there is a blank column label for race_eth)
  cols_label(
    race_eth = "",
    Dem = "DEM.",
    Rep = "REP.",
    Und = "UND."
  ) %>%
  # Adds a source note to the footer of a gt table. 
    tab_source_note(source_note = "Source: New York Times Upshot/Siena College 2018 live polls") %>% 
  # formats as a percent, and shows 0 decimal places
  fmt_percent(columns = vars(Dem, Rep, Und),
              decimals = 0) %>% 
  # turn the 0 back into an NA
  na_if(0) %>%
  # reformats the NA to a dash 
  fmt_missing(columns = vars(Und), rows = 4)
```

##Question 3
```{r graph, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 7.7}
graph <- elections %>% 
  # Select the values that are necessary
  #we are graphing final weight against different levels of education, so we need both of those values
  select(educ, final_weight) %>% 
  # Take out values where the education was not recorded
  filter(educ != "[DO NOT READ] Refused") %>% 
  # Reorders the y axis of the graph (the education levels) to match the original graph
  mutate(educ = fct_relevel(educ, "Grade school", "High school", "Some college or trade school", "Bachelors' degree", "Graduate or Professional Degree"))

# Adds a plot, with the x axis being education and the y axis the final weight 
ggplot(graph, aes(x = educ, y = final_weight)) +
  # use geom_violin to match the graph used in the figure. geom_violin is a mirrored density plot 
  geom_violin() + 
  # jitter to add better visibility of the dots. Width - adds the amount of jitter. Size - changes the size of the dots
  # alpha - changes the transparency of the dots aesthetically 
  geom_jitter(width = 0.2, size = 1.0, alpha = 0.5) + 
  # flips the coordinate values of educ and final_weight, because otherwise the violins show up as lines and not the mirroed density plots
    coord_flip() +
  # Adds the title, subtitle, and caption labels
    labs(title = "More Educated Matter Less in North Carolina 9th ",
          subtitle = "Poll gives more weight to people who are less likely to participate in polls",
          caption = "New York Times Upshot/Siena College 2018 live polls") +
  # Adds a y-axis label
    ylab("Weight Given to Respondent in Calculating Poll Results")
```

##Question 4
```{r, echo=FALSE}
# I wanted to explore the Democrat/Republican voting patterns between men and women 
# I assigned question4 to the dataset I filtered for the purposes of Question 4. 
question4 <- elections %>%
  # I selected gender and response because I was interested in seeing the relationship between these two variables. 
  select(gender, response) %>%
  # Because there are third party votes and Und votes, I filtered for Rep and Dem to filter for the two main parties
  filter(response %in% c("Rep", "Dem")) %>% 
  # I grouped by response and gender to get values for each gender's responses (i.e. how many women voted Republican, how many women voted democrat, how many men voted republican, how many men voted democrat). The first step is to group by response and gender.
  group_by(response, gender) %>% 
  # The second step is to count for each of these subsets. 
  count()

# created a graph, with x being response and y being the number in each subset. Fill is by response, so the fill is dependent on the type of response (i.e. different colors for Dem and Rep responses)
ggplot(question4, aes(x = response, y = n, fill = response)) +
  # geom_col to create a column graph
  geom_col() +
  # facet wrap by gender
  facet_wrap(~gender) +
  # label the x axis
  xlab(label = "Response by Gender") +
  # label the y axis
  ylab(label = "Number of Votes") + 
  # label the title, subtitle, and caption
  labs(title = "Number of Responses by Gender",
       subtitle = "Votes for Democratic and Republican Candidates by Gender",
       caption = "New York Times Upshot/Siena College 2018 live polls") +
  # fill the colors manually, so that democrats are blue and republicans are red
  scale_fill_manual(values=c("blue", "red"))
# there are a lot of interesting insights that you can find when you split the data up into these four subsets. First, more men voted republican than democrat, and more women voted democrat than republican. Also, based on the data, it looks like more men than woman, even though the gender split is supposed to be 50/50. This may suggest women have lower voter turnout, or some other interest points. 
```

Collaborators: Mike Silva, Jennifer Li, Neil Khurana